---
title: "Simulation Based on FARMM Data"
author: "Pixu Shi"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    theme: united
---

# Preparation

## Library

```{r}
reticulate::use_condaenv("/Users/pixushi/miniconda3/envs/qiime2-2021.2")
# for computing
library(reticulate) # run py codes
library(vegan) # distance matrix
library(phyloseq) # to handle phyloseq object
library(PERMANOVA) # permanova
library(randomForest) # random forest
library(PRROC) # roc and pr
library(tempted)
library(microTensor)

# for data
library(qiime2R) # read in Qiime artifacts
library(dplyr) # data formatting
library(readr)

# for plotting
library(ggpubr)
library(ggplot2)
library(gridExtra)
library(RColorBrewer)

rm(list=ls())
# set working directory to be where the current script is located
mydir <- dirname(rstudioapi::getSourceEditorContext()$path)
setwd(mydir)
color_RB <- brewer.pal(3,'Set1')
```

## Read in the data

```{r}
# meta data table
load("data/df_samples.RData")
meta_tab <- as.data.frame(df_samples)
rownames(meta_tab) <- meta_tab$SampleID
metauni <- unique(meta_tab[,c('SubjectID', 'study_group')])
rownames(metauni) <- metauni$SubjectID

prop_tab <- readr::read_tsv("data/kraken_results.tsv")
prop_tab <- as.data.frame(prop_tab)
rownames(prop_tab) <- prop_tab[,1]
prop_tab <- prop_tab[,-1]
prop_tab <- t(prop_tab)
prop_tab <- prop_tab[rownames(meta_tab),]
prop_tab <- prop_tab[,colSums(prop_tab!=0)>=5]
prop_tab <- prop_tab[rowSums(prop_tab!=0)>=5,]
meta_tab <- meta_tab[rownames(prop_tab),]
dim(prop_tab)
dim(meta_tab)
```



```{r setup}
depth <- c(5e4, 1e5, 5e5)
nkeep <- c(2:10)
nsim <- 100
npc <- 2
```


## Plot timeline of study

```{r timeline}
meta_reordered <- meta_tab[order(meta_tab$study_group, decreasing=T),]
meta_reordered$SubjectID <- factor(meta_reordered$SubjectID, 
                                 unique(meta_reordered$SubjectID))
p_timeline <- ggplot(data=meta_reordered, 
       aes(x=study_day, y=SubjectID, 
           group=SubjectID, color=study_group, shape=study_group)) +
  geom_line() + geom_point() + scale_color_manual(values=color_RB) +
  labs(y="Host ID", x="Study Day", color="Dietary Group", shape="Dietary Group") + 
  theme_bw() +
  theme(legend.position="bottom")
p_timeline
pdf('../figure_table/FARMM_timeline.pdf', width=7.7, height=5)
print(p_timeline)
dev.off()
```


# Run with the entire dataset

## TEMPTED

```{r run_tempted_ori}
npc <- 2

set.seed(0)
count_tab <- t(apply(prop_tab,1,function(x){rmultinom(1,50000,x)}))

subdata <- format_tempted(count_tab, meta_tab$study_day, meta_tab$SubjectID, 
                        threshold=1, pseudo=0.5,
                        transform="clr")
# run tempted with all subjects and run permanova test
svd_sub <- svd_centralize(subdata)
res_tempted <- tempted(svd_sub$datlist, r = npc, resolution = 101, smooth=1e-4)
agg_feat <- aggregate_feature(res_tempted, NULL, subdata)
aggfeat_mat <- reshape(agg_feat$metafeature_aggregate[,1:4],
                       idvar=c("subID","timepoint") ,
                       v.names=c("value"), timevar="PC",
                       direction="wide")
colnames(aggfeat_mat)[1] <- 'SubjectID'
aggfeat_mat <- merge(aggfeat_mat, metauni, by='SubjectID')
aggfeat_mat[,2+1:npc] <- apply(aggfeat_mat[,2+1:npc], 2, scale)
tempted_sub <- as.data.frame(res_tempted$A_hat)
tempted_sub$`intercept` <- svd_sub$A_tilde
tempted_sub$SubjectID <- rownames(tempted_sub)
tempted_sub <- merge(metauni, tempted_sub)
colnames(tempted_sub)[1:npc+2] <- paste0("PC",1:npc)

par(mfrow=c(1,3))
with(tempted_sub, boxplot(PC1~study_group))
with(tempted_sub, boxplot(PC2~study_group))
with(tempted_sub, plot(PC1, PC2, col=1+(study_group=="EEN")))
```

## microTensor

```{r run_microtensor_ori}
X_array <- array(NA, dim = c(ncol(count_tab),
                         length(unique(meta_tab$SubjectID)),
                         16))
dimnames(X_array) <- list(colnames(count_tab),
                          unique(meta_tab$SubjectID),
                          0:15)
for(k in seq(1, dim(X_array)[3])) {
  k_df_samples <- meta_tab %>% 
    dplyr::filter(study_day == k-1)
  k_df_samples <- k_df_samples[!duplicated(k_df_samples$SubjectID),]
  X_array[, k_df_samples$SubjectID, k] <- 
    t(count_tab[rownames(k_df_samples), ])
}
set.seed(0)
fit_microTensor <- 
  microTensor::microTensor(X = X_array, R = npc, 
                           nn_t = TRUE, ortho_m = TRUE,
                           weighted = TRUE,
                           control = list(L_init = 2,
                                          gamma = 2,
                                          maxit = 1000,
                                          verbose = TRUE))
micro_sub <- as.data.frame(fit_microTensor$s)
colnames(micro_sub) <- paste0("PC",1:npc)
micro_sub$SubjectID <- dimnames(X_array)[[2]]
micro_sub <- merge(metauni, micro_sub)

par(mfrow=c(1,3))
with(micro_sub, boxplot(PC1~study_group))
with(micro_sub, boxplot(PC2~study_group))
with(micro_sub, plot(PC1, PC2, col=1+(study_group=="EEN")))
```

## CTF

```{r run_ctf_ori}
ss <- 0
ntime <- 0
depthkk <- 0
npc <- 2
meta_sub <- meta_tab[,c("study_day", "SubjectID", "study_group")]
count_sub <- as.data.frame(count_tab)
# if a sample only has one taxon, CTF fails to run, so such samples are removed
count_sub <- count_sub[rowSums(count_sub!=0)>1,]
meta_sub <- meta_sub[rownames(count_sub),]
meta_sub$SubjectID <- as.character(meta_sub$SubjectID)
metauni <- unique(meta_sub[,c("SubjectID", "study_group")])
py_run_file(file="run_farmm_ctf.py",convert=F)

metauni_sub <- unique(meta_sub[,c('SubjectID', 'study_group')])
rownames(metauni_sub) <- metauni_sub$SubjectID
fname <- paste0("subject-biplot_sim", ss, "_ntime", ntime, "_depth", depthkk, "K.qza")
ctf_sub <- read_qza(file.path("simresult_ctf", fname))$data$Vectors
colnames(ctf_sub)[1] <- "SubjectID"
ctf_sub <- merge(ctf_sub, metauni_sub)

par(mfrow=c(1,3))
with(ctf_sub, boxplot(PC1~study_group))
with(ctf_sub, boxplot(PC2~study_group))
with(ctf_sub, plot(PC1, PC2, col=1+(study_group=="EEN")))
```


# Simulated data


```{r simulate_data, eval=FALSE}
meta_tab2 <- meta_tab[,c("study_day", "SubjectID", "study_group")]
set.seed(0)
for (ss in 1:nsim){
  print(ss)
  for (jj in 1:length(nkeep)){
    sampleID_temp <- NULL
    for (i in 1:nrow(metauni)){
      meta_sub <- dplyr::filter(meta_tab2, SubjectID==metauni$SubjectID[i])
      meta_sub <- dplyr::arrange(meta_sub, -study_day)
      meta_sub <- meta_sub[!duplicated(meta_sub$study_day),]
      meta_sub <- dplyr::arrange(meta_sub, study_day)
      sel <- sample(rownames(meta_sub), 
                    size=min(nrow(meta_sub), nkeep[jj]))
      sel <- sel[order(meta_sub[sel,'study_day'])]
      sampleID_temp <- c(sampleID_temp, sel)
    }
    meta_sub <- meta_tab2[sampleID_temp,]
    for (kk in 1:length(depth)){
          count_sub <- t(apply(prop_tab[sampleID_temp,], 1, 
                               function(x){rmultinom(1,50000,x)}))
          write.csv(cbind(meta_sub, count_sub), file=paste0("simdata/realsim", ss, "_ntime", nkeep[jj], "_depth", depth[kk]/1000, "K.csv"))
    }
  }
}
```




# TEMPTED

## Run TEMPTED

Save subject loading and trajectory

```{r run_tempted, eval=FALSE}
for (ss in 1:nsim){
  print(ss)
  for (jj in 1:length(nkeep)){
    for (kk in 1:length(depth)){
      dat_sub <- read.csv(paste0("simdata/realsim", ss, "_ntime", nkeep[jj], "_depth", depth[kk]/1000, "K.csv"), header=T, row.names=1)
      count_sub <- dat_sub[,-c(1:3)]
      meta_sub <- dat_sub[,1:3]
      meta_sub$SubjectID <- as.character(meta_sub$SubjectID)
      metauni <- unique(meta_sub[,c("SubjectID", "study_group")])
      subdata <- format_tempted(count_sub, meta_sub$study_day, meta_sub$SubjectID, 
                              threshold=1, pseudo=0.5,
                              transform="clr")
      # run tempted with all subjects
      svd_sub <- svd_centralize(subdata)
      res_tempted <- tempted(svd_sub$datlist, r = npc, resolution = 101, smooth=0.0001)
      # plot_time_loading(res_tempted)
      agg_feat <- aggregate_feature(res_tempted, NULL, subdata)
      aggfeat_mat <- reshape(agg_feat$metafeature_aggregate[,1:4],
                             idvar=c("subID","timepoint") ,
                             v.names=c("value"), timevar="PC",
                             direction="wide")
      colnames(aggfeat_mat)[1] <- 'SubjectID'
      aggfeat_mat <- merge(aggfeat_mat, metauni, by='SubjectID')
      aggfeat_mat[,2+1:npc] <- apply(aggfeat_mat[,2+1:npc], 2, scale)
      tempted_sub <- as.data.frame(res_tempted$A_hat)
      tempted_sub$`intercept` <- svd_sub$A_tilde
      tempted_sub$SubjectID <- rownames(tempted_sub)
      tempted_sub <- merge(metauni, tempted_sub)
      # with(tempted_sub, plot(`Component 1`, `Component 3`, col=as.numeric(as.factor(study_group))+1))
      # with(tempted_sub, boxplot(`Component 3`~study_group))
      write.csv(aggfeat_mat, 
                file=paste0('simresult_tempted/tempted_traj_sim',ss,'_ntime',nkeep[jj], "_depth", depth[kk]/1000, 'K.csv'))
      write.csv(tempted_sub, 
                file=paste0('simresult_tempted/tempted_subj_sim',ss,'_ntime',nkeep[jj], "_depth", depth[kk]/1000, 'K.csv'))
    }
  }
}
```

## In-sample ROC

-   Random forest
-   Logistic regression

```{r tempted_subj_insample_auc, eval=FALSE}
for (kk in 1:length(depth)){
  roc_sub_rf_tempted <- matrix(NA, nsim, length(nkeep))
  colnames(roc_sub_rf_tempted) <- paste0("nsample", nkeep)
  roc_sub_glm_tempted <- roc_sub_rf_tempted
  for (jj in 1:length(nkeep)){
    print(jj)
    for (ss in 1:nsim){
      fname <- paste0('tempted_subj_sim',ss,'_ntime',nkeep[jj], '_depth', depth[kk]/1000, 'K.csv')
      tempted_sub <- read.csv(file.path("simresult_tempted", fname), row.names=1)
      # random forest
      rf_fit <- randomForest(as.factor(study_group) ~ Component.1+Component.2,
                     data = tempted_sub)
      predprob_rf <- predict(rf_fit, type = "prob")[,"EEN"]
      ind_een <- tempted_sub$study_group=="EEN"
      roc_sub_rf_tempted[ss,jj] <- roc.curve(predprob_rf[ind_een], 
                                   predprob_rf[!ind_een])$auc
      # logistic regression
      glm_fit <- glm(I(study_group=="EEN") ~ Component.1+Component.2, 
                    family = binomial(link="logit"),
                    data = tempted_sub)
      predprob_glm <- predict(glm_fit, type = "response")
      ind_een <- tempted_sub$study_group=="EEN"
      roc_sub_glm_tempted[ss,jj] <- roc.curve(predprob_glm[ind_een], 
                                        predprob_glm[!ind_een])$auc
    }
  }
  write.csv(roc_sub_rf_tempted, 
            file=paste0('result/realsim_farmm_roc_sub_rf_tempted_clr_', depth[kk]/1000, 'K.csv'))
  write.csv(roc_sub_glm_tempted, 
            file=paste0('result/realsim_farmm_roc_sub_glm_tempted_clr_', depth[kk]/1000, 'K.csv'))
}
```

## Out-of-Sample 

-   Random forest
-   Logistic regression

```{r run_oos, eval=FALSE}
for (kk in 1:length(depth)){
  roc_glm_oos <- matrix(NA, nsim, length(nkeep))
  colnames(roc_glm_oos) <- paste0("nsample", nkeep)
  pr_rf_oos <- roc_rf_oos <- pr_glm_oos <- roc_glm_oos
  for (jj in 1:length(nkeep)){
    print(jj)
    for (ss in 1:nsim){
      dat_sub <- read.csv(paste0("simdata/realsim", ss, "_ntime", nkeep[jj], "_depth", depth[kk]/1000, "K.csv"), header=T, row.names=1)
      count_sub <- dat_sub[,-c(1:3)]
      meta_sub <- dat_sub[,1:3]
      meta_sub$SubjectID <- as.character(meta_sub$SubjectID)
      subdata <- format_tempted(count_sub, meta_sub$study_day, meta_sub$SubjectID, 
                              threshold=1, pseudo=0.5,
                              transform="clr")
      metauni_sub <- unique(meta_sub[,c("SubjectID", "study_group")])
      metauni_sub$study_group_EEN <- metauni_sub$study_group=="EEN"
      # leave one out prediction
      predprob_glm <- predprob_rf <- rep(NA, length(subdata))
      
      for (ii in 1:length(subdata)){
        print(ii)
        svd_train <- svd_centralize(subdata[-ii])
        res_train <- tempted(svd_train$datlist, r = npc, resolution = 101, smooth=0.0001)
        A_test <- est_test_subject(subdata[ii], res_train, svd_train)
        dftrain <- data.frame(y=metauni_sub[-ii,'study_group_EEN'], x=res_train$A)
        dftest <- data.frame(y=metauni_sub[ii,'study_group_EEN'], x=A_test)
        # logistic regression
        glm_fit <- glm(y ~ ., data = dftrain, family = "binomial")
        predprob_glm[ii] <- predict(glm_fit, newdata=dftest, type = "response")
        # random forest
        rf_fit <- randomForest(as.factor(y) ~ ., data = dftrain)
        predprob_rf[ii] <- predict(rf_fit, newdata=dftest, type = "prob")[,"TRUE"]
      }
      roc_rf_oos[ss,jj] <- roc.curve(predprob_rf[metauni_sub$study_group_EEN], 
                                          predprob_rf[!metauni_sub$study_group_EEN])$auc
      roc_glm_oos[ss,jj] <- roc.curve(predprob_glm[metauni_sub$study_group_EEN], 
                                          predprob_glm[!metauni_sub$study_group_EEN])$auc
    }
    write.csv(roc_rf_oos, file=paste0("result/realsim_farmm_roc_oos_rf_tempted_", depth[kk]/1000, "K.csv"))
    write.csv(roc_glm_oos, file=paste0("result/realsim_farmm_roc_oos_glm_tempted_", depth[kk]/1000, "K.csv"))
  }
}

```

# CTF

## Run CTF

Save subject distance matrix and trajectory

```{r run_ctf, eval=FALSE}
for (ss in 1:nsim){
  print(ss)
  set.seed(0)
  for (jj in 1:length(nkeep)){
    ntime <- nkeep[jj]
    for(kk in 1:length(depth)){
      depthkk <- depth[kk]/1000
      dat_sub <- read.csv(paste0("simdata/realsim", ss, "_ntime", nkeep[jj], '_depth', depth[kk]/1000, "K.csv"), header=T, row.names=1)
      meta_sub <- dat_sub[,1:3]
      count_sub <- dat_sub[,-c(1:3)]
      # if a sample only has one taxon, CTF fails to run, so such samples are removed
      count_sub <- count_sub[rowSums(count_sub!=0)>1,]
      meta_sub <- meta_sub[rownames(count_sub),]
      meta_sub$SubjectID <- as.character(meta_sub$SubjectID)
      metauni <- unique(meta_sub[,c("SubjectID", "study_group")])
      py_run_file(file="run_farmm_ctf.py",convert=F)
    }
  }
}

```

## In-sample ROC

-   Random forest
-   Logistic regression

```{r ctf_subj_insample_auc, eval=FALSE}
for (kk in 1:length(depth)){
  roc_sub_rf_ctf <- matrix(NA, nsim, length(nkeep))
  colnames(roc_sub_rf_ctf) <- paste0("nsample", nkeep)
  roc_sub_glm_ctf <- roc_sub_rf_ctf
  for (jj in 1:length(nkeep)){
    print(jj)
    ntime <- nkeep[jj]
    for (ss in 1:nsim){
      meta_sub <- read.csv(paste0("simdata/realsim", ss, "_ntime", nkeep[jj], "_depth", depth[kk]/1000, "K.csv"), header=T, row.names=1)
      meta_sub$SubjectID <- as.character(meta_sub$SubjectID)
      metauni_sub <- unique(meta_sub[,c('SubjectID', 'study_group')])
      rownames(metauni_sub) <- metauni_sub$SubjectID
      fname <- paste0("subject-biplot_sim", ss, "_ntime", ntime, "_depth", depth[kk]/1000, "K.qza")
      ctf_sub <- read_qza(file.path("simresult_ctf", fname))$data$Vectors
      colnames(ctf_sub)[1] <- "SubjectID"
      ctf_sub <- merge(ctf_sub, metauni_sub)
      # random forest
      rf_fit <- randomForest(as.factor(study_group)~PC1+PC2, data=ctf_sub)
      predprob_rf <- predict(rf_fit, type = "prob")[,"EEN"]
      ind_een <- ctf_sub$study_group=="EEN"
      roc_sub_rf_ctf[ss,jj] <- roc.curve(predprob_rf[ind_een], 
                                    predprob_rf[!ind_een])$auc
      # logistic regression
      glm_fit <- glm(I(study_group=="EEN") ~ PC1+PC2, 
                    family = binomial(link="logit"),
                    data = ctf_sub)
      predprob_glm <- predict(glm_fit, type = "response")
      ind_een <- ctf_sub$study_group=="EEN"
      roc_sub_glm_ctf[ss,jj] <- roc.curve(predprob_glm[ind_een], 
                                        predprob_glm[!ind_een])$auc
    }
  }
  write.csv(roc_sub_rf_ctf, 
            file=paste0('result/realsim_farmm_roc_sub_rf_ctf_', depth[kk]/1000, 'K.csv'))
  write.csv(roc_sub_glm_ctf, 
            file=paste0('result/realsim_farmm_roc_sub_glm_ctf_', depth[kk]/1000, 'K.csv'))
}
```


# microTensor

## Run microTensor

Save subject loading and trajectory

```{r run_microtensor, eval=FALSE}
outdir <- "simresult_microtensor"
for (ss in 1:nsim){
  print(ss)
  for (jj in 1:length(nkeep)){
    for (kk in 1:length(depth)){
      dat_sub <- read.csv(paste0("simdata/realsim", ss, "_ntime", nkeep[jj], "_depth", depth[kk]/1000, "K.csv"), header=T, row.names=1)
      count_sub <- dat_sub[,-c(1:3)]
      count_sub <- count_sub[,colMeans(count_sub==0)<=0.95]
      meta_sub <- dat_sub[,1:3]
      meta_sub$SubjectID <- as.character(meta_sub$SubjectID)
      meta_sub$sampleid <- rownames(meta_sub)
      metauni <- unique(meta_sub[,c("SubjectID", "study_group")])
      set.seed(0)
      X_array <- array(NA, dim = c(ncol(count_sub),
                               length(unique(meta_sub$SubjectID)),
                               16))
      dimnames(X_array) <- list(colnames(count_sub),
                                unique(meta_sub$SubjectID),
                                0:15)
      for(k in seq(1, dim(X_array)[3])) {
        k_df_samples <- meta_sub %>% 
          dplyr::filter(study_day == k-1)
        k_df_samples <- k_df_samples[!duplicated(k_df_samples$SubjectID),]
        X_array[, k_df_samples$SubjectID, k] <- 
          t(count_sub[rownames(k_df_samples), ])
      }
      set.seed(1)
      fit_microTensor <- 
        microTensor::microTensor(X = X_array, R = npc, 
                                 nn_t = TRUE, ortho_m = TRUE,
                                 weighted = TRUE)
      micro_sub <- as.data.frame(fit_microTensor$s)
      colnames(micro_sub) <- paste0("PC",1:npc)
      micro_sub$SubjectID <- dimnames(X_array)[[2]]
      micro_sub <- merge(metauni, micro_sub)
      write.csv(micro_sub, 
                file=paste0(outdir, "/micro_subj_sim", ss, "_ntime", nkeep[jj], "_depth", depth[kk]/1000, "K.csv"))
      micro_traj <- microTensor::create_loading(fit_decomp = fit_microTensor,
                                                feature_names = dimnames(X_array)[[1]],
                                                subject_names = dimnames(X_array)[[2]],
                                                time_names = 0:15,
                                                class = "sample")
      colnames(micro_traj) <- c("SubjectID", "study_day", paste0("PC",1:npc))
      micro_traj <- merge(meta_sub[,c("sampleid","study_day","SubjectID", "study_group")], micro_traj)
      write.csv(micro_traj, 
                file=paste0(outdir, "/micro_traj_sim", ss, "_ntime", nkeep[jj], "_depth", depth[kk]/1000, "K.csv"))
    }
  }
}
```

## In-sample ROC

-   Random forest
-   Logistic regression

```{r micro_subj_insample_auc, eval=FALSE}
suffix <- "microtensor"
for (kk in 1:length(depth)){
  roc_sub_rf_micro <- matrix(NA, nsim, length(nkeep))
  colnames(roc_sub_rf_micro) <- paste0("nsample", nkeep)
  roc_sub_glm_micro <- roc_sub_rf_micro
  for (jj in 1:length(nkeep)){
    print(jj)
    for (ss in 1:nsim){
      fname <- paste0('micro_subj_sim',ss,'_ntime',nkeep[jj], '_depth', depth[kk]/1000, 'K.csv')
      micro_sub <- read.csv(file.path(paste0("simresult_", suffix), fname), row.names=1)
      # random forest
      rf_fit <- randomForest(as.factor(study_group) ~ PC1+PC2,
                     data = micro_sub)
      predprob_rf <- predict(rf_fit, type = "prob")[,"EEN"]
      ind_een <- micro_sub$study_group=="EEN"
      roc_sub_rf_micro[ss,jj] <- roc.curve(predprob_rf[ind_een], 
                                        predprob_rf[!ind_een])$auc
      # logistic regression
      glm_fit <- glm(I(study_group=="EEN") ~ PC1+PC2, 
                    family = binomial(link="logit"),
                    data = micro_sub)
      predprob_glm <- predict(glm_fit, type = "response")
      ind_een <- micro_sub$study_group=="EEN"
      roc_sub_glm_micro[ss,jj] <- roc.curve(predprob_glm[ind_een], 
                                        predprob_glm[!ind_een])$auc
    }
  }
  write.csv(roc_sub_rf_micro, 
          file=paste0('result/realsim_farmm_roc_sub_rf_', suffix, '_', depth[kk]/1000, 'K.csv'))
  write.csv(roc_sub_glm_micro, 
          file=paste0('result/realsim_farmm_roc_sub_glm_', suffix, '_', depth[kk]/1000, 'K.csv'))
}
```


# Summarize results

## Summarize subject-level ROC and PR

```{r summary_sub_classification}
method  <- c("tempted_clr", "ctf", "microtensor")
measure <- c("roc")
classify <- c("rf", "glm")

tab_auc_subj <- NULL

for (ms in measure){
  for (cls in classify){ 
    # in sample
    for (mthd in method){
      for (kk in 1:length(depth)){
        fname <- paste0("result/realsim_farmm_", ms, "_sub_", cls, "_", mthd, "_", depth[kk]/1000, "K.csv")
        tab0 <- read.csv(fname, row.names=1)
        tab0 <- 1-as.matrix(tab0)
        tab_tmp <- data.frame(sim=1:nsim,
                        auc=as.vector(tab0),
                        nsample=as.vector(t(matrix(rep(nkeep,nsim),length(nkeep),nsim))),
                        method=toupper(mthd), measure=toupper(ms),
                        classify=cls,
                        type="In-Sample", depth=depth[kk])
        tab_auc_subj <- rbind(tab_auc_subj, tab_tmp)
        # out of sample
        fname <- paste0("result/realsim_farmm_", ms, "_oos_", cls, "_tempted_", depth[kk]/1000, "K.csv")
        tab0 <- read.csv(fname, row.names=1)
        tab0 <- 1-as.matrix(tab0)
        tab_tmp <- data.frame(sim=1:nsim,
                              auc=as.vector(tab0),
                        nsample=as.vector(t(matrix(rep(nkeep,nsim),length(nkeep),nsim))),
                        method="TEMPTED", measure=toupper(ms), classify=cls,
                        type="Out-of-Sample", depth=depth[kk])
        tab_auc_subj <- rbind(tab_auc_subj, tab_tmp)
      }
    }
  }  
}
tab_auc_subj$classify <- gsub("rf", "Random Forest", tab_auc_subj$classify)
tab_auc_subj$classify <- gsub("glm", "Logistic Regression", tab_auc_subj$classify)
tab_auc_subj$method <- gsub("TEMPTED_CLR", "TEMPTED", tab_auc_subj$method)
tab_auc_subj$method <- gsub("MICROTENSOR", "microTensor", tab_auc_subj$method)
tab_auc_subj$depth_name <- paste0("Reads per Sample = ", 
                                  tab_auc_subj$depth/1000, "K")
tab_auc_subj$depth_name <- factor(tab_auc_subj$depth_name, 
                                     levels=paste0("Reads per Sample = ", 
                                  depth/1000, "K"))

# summarize mean and error band for each method
tab_auc_subj$nsample <- as.factor(tab_auc_subj$nsample)
tab1 <- aggregate(auc~nsample+measure+method+classify+type+depth_name, data=tab_auc_subj, 
                  FUN=mean)
names(tab1)[7] <- 'mean'
tab2 <- aggregate(auc~nsample+measure+method+classify+type+depth_name, data=tab_auc_subj, 
                  FUN=function(x){sd(x)/sqrt(length(x))})
names(tab2)[7] <- 'se'
rownames(tab1) <-rownames(tab2) <- NULL
tab_auc_subj_summary <- merge(tab1, tab2)
tab_auc_subj_summary$nsample <- factor(tab_auc_subj_summary$nsample, 
                                     level=as.character(nkeep))


# get the best AUC of the two methods
tab_auc_subj_best <- aggregate(auc~sim+nsample+measure+method+type+depth_name, data=tab_auc_subj, 
                  FUN=min)

# summarize mean and error band for the best of two methods
tab_auc_subj_best$nsample <- as.factor(tab_auc_subj_best$nsample)
tab1 <- aggregate(auc~nsample+measure+method+type+depth_name, data=tab_auc_subj_best, 
                  FUN=mean)
names(tab1)[6] <- 'mean'
tab2 <- aggregate(auc~nsample+measure+method+type+depth_name, data=tab_auc_subj_best, 
                  FUN=function(x){sd(x)/sqrt(length(x))})
names(tab2)[6] <- 'se'
rownames(tab1) <-rownames(tab2) <- NULL
tab_auc_subj_best_summary <- merge(tab1, tab2)
tab_auc_subj_best_summary$nsample <- factor(tab_auc_subj_best_summary$nsample, 
                                     level=as.character(nkeep))
```

## Plot all subject-level AUC-ROC error

### Logistic regression and random forest separately

```{r plot_sub_auc}
ann_text <- tab_auc_subj_summary[2,]
ann_text$mean <- 0.515
ann_text$nsample <- 7
color_method <- c('#33a02c', #green for CTF
                  '#0096FF', #blue for microTensor
                  '#5A5A5A') #gray for TEMPTED

p_subj_roc <- ggplot(data=dplyr::filter(tab_auc_subj_summary, measure=="ROC"), 
                           aes(x=nsample, y=mean, group=paste0(type,method), color=method)) + 
  geom_line(size=1, position=position_dodge(0.5), aes(linetype=type)) + 
  geom_point(size=2, position=position_dodge(0.5)) +
  geom_errorbar(aes(ymin=mean-2*se, ymax=mean+2*se, width=0.5), 
                position=position_dodge(0.5), size=1) + 
  geom_hline(yintercept=0.5, linetype="dotted", color = "black", size=1) +
  geom_text(data = ann_text,label = "At random", color="black") +
  facet_grid(classify~depth_name) +
  labs(y='AUC-ROC error', x='Number of time points') + 
  ggtitle("Subject Level") +
  scale_color_manual(values=color_method) +
  theme_bw()
p_subj_roc
```

### Smallest of the two classification methods

```{r}
ann_text <- tab_auc_subj_summary[2,]
ann_text$mean <- 0.515
ann_text$nsample <- 7
color_method <- c('#33a02c', #green for CTF
                  '#0096FF', #blue for microTensor
                  '#5A5A5A') #gray for TEMPTED

p_subj_roc_best <- ggplot(data=dplyr::filter(tab_auc_subj_best_summary, measure=="ROC"), 
                           aes(x=nsample, y=mean, group=paste0(type,method), color=method)) + 
  geom_line(size=1, position=position_dodge(0.5), aes(linetype=type)) + 
  geom_point(size=2, position=position_dodge(0.5)) +
  geom_errorbar(aes(ymin=mean-2*se, ymax=mean+2*se, width=0.5), 
                position=position_dodge(0.5), size=1) + 
  geom_hline(yintercept=0.5, linetype="dotted", color = "black", size=1) +
  geom_text(data = ann_text,label = "At random", color="black") +
  facet_wrap(.~depth_name) +
  labs(y='AUC-ROC error', x='Number of time points') + 
  ggtitle("Subject Level") +
  scale_color_manual(values=color_method) +
  theme_bw() + 
  theme(legend.position="none")
p_subj_roc_best
```


## Print plot for manuscript

```{r plot_all}
ggsave('../figure_table/realsim_farmm.pdf', width=10, height=4, plot=p_subj_roc_best)
```
