---
title: "Simulation Based on ECAM Data for More Methods"
author: "Pixu Shi"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    theme: united
---

# Preparation

## Library

```{r}
reticulate::use_condaenv("/Users/pixushi/miniconda3/envs/qiime2-2021.2")
rm(list=ls())
# for computing
library(reticulate) # run py codes
library(vegan) # distance matrix
library(PERMANOVA) # permanova
library(randomForest) # random forest
library(PRROC) # roc and pr
library(tempted)
library(microTensor)

# for data
library(qiime2R) # read in Qiime artifacts
library(dplyr) # data formatting

# for plotting
library(ggpubr)
library(ggplot2)
library(gridExtra)
library(RColorBrewer)

# set working directory to be where the current script is located
mydir <- dirname(rstudioapi::getSourceEditorContext()$path)
setwd(mydir)
color_RB <- brewer.pal(3,"Set1")[1:2]
```

## Read in the data

```{r read_data}
count_tab <- read.csv("data/otu_count_cleaned_q2.csv", row.names=1)
meta_tab <- read.csv("data/otu_metadata_cleaned_q2.csv", row.names=1)
taxon_tab <- read.csv("data/otu_taxonomy_cleaned_q2.csv", row.name=1)
table(rownames(count_tab)==rownames(meta_tab))
meta_tab$delivery_ind <- meta_tab$delivery=="Vaginal"
# remove distal gut samples
meta_tab <- meta_tab %>%
  dplyr::filter(qiita_empo_3=="anthropogenic sample" | is.na(qiita_empo_3))
# remove subject with <9 samples
sub_rm <- names(which(table(meta_tab$studyid)<9))
meta_tab <- meta_tab %>%
  dplyr::filter(!studyid %in% sub_rm)
table(meta_tab$studyid)
count_tab <- count_tab[rownames(meta_tab),]
table(rownames(count_tab)==rownames(meta_tab))
metauni <- unique(meta_tab[,c("studyid", "delivery_ind", "diet")])
rownames(metauni) <- metauni$studyid
```

## Plot timeline of study

```{r timeline}
meta_reordered <- meta_tab[order(meta_tab$delivery, meta_tab$studyid, decreasing=T),]
meta_reordered$studyid <- factor(meta_reordered$studyid, 
                                 unique(meta_reordered$studyid))
colnames(meta_reordered)[10] <- "Delivery"
p_timeline <- ggplot(data=meta_reordered, 
       aes(x=day_of_life, y=studyid, 
           group=studyid, color=Delivery, shape=Delivery)) +
  geom_line() + geom_point() + scale_color_manual(values=color_RB) +
  labs(y="Host ID", x="Day of Life") + 
  theme_bw() +
  theme(legend.position="bottom")
p_timeline
pdf("../figure_table/ECAM_timeline.pdf", width=7.7, height=5)
print(p_timeline)
dev.off()
```

# Simulated data by resampling ECAM samples

```{r setup}
nkeep <- c(2,3,4,5,6,7,8,9)
nsim <- 100
npc <- 2
```


```{r simulate_data, eval=FALSE}
meta_tab2 <- meta_tab[,c("day_of_life", "studyid", "delivery")]
set.seed(0)
for (ss in 1:nsim){
  print(ss)
  for (jj in 1:length(nkeep)){
    sampleID_temp <- NULL
    visit <- NULL
    for (i in 1:nrow(metauni)){
      meta_sub <- dplyr::filter(meta_tab, studyid==metauni$studyid[i])
      meta_sub <- dplyr::arrange(meta_sub, -day_of_life)
      meta_sub <- meta_sub[!duplicated(meta_sub$day_of_life),]
      meta_sub <- dplyr::arrange(meta_sub, day_of_life)
      nsample1 <- sum(meta_sub$day_of_life<=365)
      nsample2 <- sum(meta_sub$day_of_life>365)
      prob_vec <- c(rep(1,nsample1), rep(2,nsample2))
      prob_vec <- prob_vec/sum(prob_vec)
      sel <- sample(rownames(meta_sub), 
                    size=min(nsample1+nsample2, nkeep[jj]),
                    prob=prob_vec)
      sel <- sel[order(meta_sub[sel,"day_of_life"])]
      sampleID_temp <- c(sampleID_temp, sel)
      visit_tmp <- 1:nkeep[jj]
      visit <- c(visit, visit_tmp)
    }
    meta_sub <- meta_tab2[sampleID_temp,]
    meta_sub$visit <- visit
    write.csv(meta_sub, file=paste0("simdata/realsim", ss, "_ntime", nkeep[jj], ".csv"))
  }
}
```

Save into csv file for the number of time points 8 and 9

```{r simulate_data_save, eval=FALSE}
for (ss in 1:nsim){
  print(ss)
  for (jj in 5:length(nkeep)){
    meta_sub <- read.csv(paste0("simdata/realsim", ss, "_ntime", nkeep[jj], ".csv"), header=T, row.names=1)
    count_sub <- count_tab[rownames(meta_sub),]
    write.csv(cbind(meta_sub, count_sub), file=paste0("simdata_upload/realsim", ss, "_ntime", nkeep[jj], ".csv"))
  }
}
```




# TEMPTED

## Run TEMPTED

Save subject loading and trajectory

```{r run_tempted, eval=FALSE}
smooth <- c(1, 0.1, 0.01, 0.005, rep(1e-4, 5))
for (ss in 1:nsim){
  print(ss)
  for (jj in 1:length(nkeep)){
    meta_sub <- read.csv(paste0("simdata/realsim", ss, "_ntime", nkeep[jj], ".csv"), header=T, row.names=1)
    count_sub <- count_tab[rownames(meta_sub),]
    meta_sub$studyid <- as.character(meta_sub$studyid)
    meta_sub$delivery_ind <- meta_sub$delivery=="Vaginal"
    metauni <- unique(meta_sub[,c("studyid", "delivery_ind")])
    subdata <- format_tempted(count_sub, meta_sub$day_of_life, meta_sub$studyid, 
                            threshold=0.95, pseudo=0.5,
                            transform="clr")
    # run tempted with all subjects and run permanova test
    svd_sub <- svd_centralize(subdata)
    res_tempted <- tempted(svd_sub$datlist, r = npc, 
                           resolution = (nkeep[jj]+16)*2, smooth=smooth[jj])
    agg_feat <- aggregate_feature(res_tempted, NULL, subdata)
    aggfeat_mat <- reshape(agg_feat$metafeature_aggregate[,1:4],
                           idvar=c("subID","timepoint") ,
                           v.names=c("value"), timevar="PC",
                           direction="wide")
    colnames(aggfeat_mat)[1] <- "studyid"
    aggfeat_mat <- merge(aggfeat_mat, metauni, by="studyid")
    aggfeat_mat[,2+1:npc] <- apply(aggfeat_mat[,2+1:npc], 2, scale)
    tempted_sub <- as.data.frame(res_tempted$A_hat)
    tempted_sub$`intercept` <- svd_sub$A_tilde
    tempted_sub$studyid <- rownames(tempted_sub)
    tempted_sub <- merge(metauni, tempted_sub)
    write.csv(aggfeat_mat, 
              file=paste0("simresult_tempted/tempted_traj_sim",ss,"_ntime",nkeep[jj], ".csv"))
    write.csv(tempted_sub, 
              file=paste0("simresult_tempted/tempted_subj_sim",ss,"_ntime",nkeep[jj], ".csv"))
  }
}
```

## Sample-level

### PERMANOVA F-value

```{r tempted_sample_Fvalue, eval=FALSE}
method <- "tempted"
Fmodel_tempted <- matrix(NA, nsim, length(nkeep))
colnames(Fmodel_tempted) <- paste0("nsample", nkeep)
for (ss in 1:nsim){
  print(ss)
  for (jj in 1:length(nkeep)){
    fname <- paste0(method,"_traj_sim",ss,"_ntime",nkeep[jj], ".csv")
    aggfeat_mat <- read.csv(file.path(paste0("simresult_", method), fname), row.names=1)
    # calculate PERMANOVA F
    dist_aggft <- vegdist(aggfeat_mat[,2+1:npc], method="euclidean")
    res_perm <- adonis2(dist_aggft ~ aggfeat_mat$delivery)
    Fmodel_tempted[ss,jj] <- res_perm$F[1]
  }
}

write.csv(Fmodel_tempted, 
          file=paste0("result/realsim_ecam_Fvalue_", method, ".csv"))
```

## Subject-level

### In-sample ROC & PR

-   logistic regression
-   random forest

```{r tempted_subj_insample_auc, eval=FALSE}
method <- "tempted"
roc_sub_glm_tempted <- matrix(NA, nsim, length(nkeep))
colnames(roc_sub_glm_tempted) <- paste0("nsample", nkeep)
pr_sub_rf_tempted <- roc_sub_rf_tempted <- pr_sub_glm_tempted <- roc_sub_glm_tempted
for (ss in 1:nsim){
  print(ss)
  for (jj in 1:length(nkeep)){
    fname <- paste0(method, "_subj_sim",ss,"_ntime",nkeep[jj], ".csv")
    tempted_sub <- read.csv(file.path("simresult_tempted", fname), row.names=1)
    # glm
    predprob_glm <- rep(NA, nrow(tempted_sub))
    for (ii in 1:nrow(tempted_sub)){
      glm_fit <- glm(delivery_ind ~ PC1+PC2,
                   data = tempted_sub[-ii,], family = "binomial")
      predprob_glm[ii] <- predict(glm_fit, newdata=tempted_sub[ii,], type = "response")
    }
    roc_sub_glm_tempted[ss,jj] <- roc.curve(predprob_glm[tempted_sub$delivery_ind], 
                                        predprob_glm[!tempted_sub$delivery_ind])$auc
    pr_sub_glm_tempted[ss,jj] <- pr.curve(predprob_glm[tempted_sub$delivery_ind], 
                                      predprob_glm[!tempted_sub$delivery_ind])$auc.integral
    # random forest
    set.seed(0)
    rf_fit <- randomForest(as.factor(delivery_ind) ~ PC1+PC2,
                   data = tempted_sub, strata=as.factor(delivery_ind))
    predprob_rf <- predict(rf_fit, type = "prob")[,"TRUE"]
    roc_sub_rf_tempted[ss,jj] <- roc.curve(predprob_rf[tempted_sub$delivery_ind], 
                                        predprob_rf[!tempted_sub$delivery_ind])$auc
    pr_sub_rf_tempted[ss,jj] <- pr.curve(predprob_rf[tempted_sub$delivery_ind], 
                              predprob_rf[!tempted_sub$delivery_ind])$auc.integral
  }
}

write.csv(roc_sub_glm_tempted, 
          file=paste0("result/realsim_ecam_roc_sub_glm_", method, ".csv"))
write.csv(pr_sub_glm_tempted, 
          file=paste0("result/realsim_ecam_pr_sub_glm_", method, ".csv"))
write.csv(roc_sub_rf_tempted, 
          file=paste0("result/realsim_ecam_roc_sub_rf_", method, ".csv"))
write.csv(pr_sub_rf_tempted, 
          file=paste0("result/realsim_ecam_pr_sub_rf_", method, ".csv"))
```

### Out-of-Sample ROC & PR

-   Logistic regression
-   Random forest

```{r run_tempted_oos, eval=FALSE}
method <- "tempted"
roc_glm_oos_tempted <- matrix(NA, nsim, length(nkeep))
colnames(roc_glm_oos_tempted) <- paste0("nsample", nkeep)
pr_rf_oos_tempted <- roc_rf_oos_tempted <- 
  pr_glm_oos_tempted <- roc_glm_oos_tempted
smooth <- c(1, 0.1, 0.01, 0.005, rep(1e-4, 5))
for (ss in 1:nsim){
  print(ss)
  for (jj in 1:length(nkeep)){
    meta_sub <- read.csv(paste0("simdata/realsim", ss, "_ntime", nkeep[jj], ".csv"), header=T, row.names=1)
    count_sub <- count_tab[rownames(meta_sub),]
    meta_sub$studyid <- as.character(meta_sub$studyid)
    meta_sub$delivery_ind <- meta_sub$delivery=="Vaginal"
    subdata <- format_tempted(count_sub, meta_sub$day_of_life, meta_sub$studyid, 
                            threshold=0.95, pseudo=0.5, transform="clr")
    metauni_sub <- unique(meta_sub[,c("studyid", "delivery_ind")])
    # leave one out prediction
    predprob_glm <- predprob_rf <- rep(NA, length(subdata))
    
    for (ii in 1:length(subdata)){
      print(ii)
      svd_train <- svd_centralize(subdata[-ii])
      res_train <- tempted(svd_train$datlist, r = npc, 
                           resolution = (nkeep[jj]+16)*2, smooth=smooth[jj])
      A_test <- est_test_subject(subdata[ii], res_train, svd_train)
      dftrain <- data.frame(y=metauni[-ii,"delivery_ind"], x=res_train$A)
      dftest <- data.frame(y=metauni[ii,"delivery_ind"], x=A_test)
      # logistic regression
      glm_fit <- glm(y ~ ., data = dftrain, family = "binomial")
      predprob_glm[ii] <- predict(glm_fit, newdata=dftest, type = "response")
      # random forest
      rf_fit <- randomForest(as.factor(y) ~ ., data = dftrain,
                             strata=as.factor(y))
      predprob_rf[ii] <- predict(rf_fit, newdata=dftest, type = "prob")[,"TRUE"]
    }
    roc_glm_oos_tempted[ss,jj] <- roc.curve(predprob_glm[metauni_sub$delivery_ind], 
                                        predprob_glm[!metauni_sub$delivery_ind])$auc
    pr_glm_oos_tempted[ss,jj] <- pr.curve(predprob_glm[metauni_sub$delivery_ind], 
                                      predprob_glm[!metauni_sub$delivery_ind])$auc.integral
    roc_rf_oos_tempted[ss,jj] <- roc.curve(predprob_rf[metauni_sub$delivery_ind], 
                                        predprob_rf[!metauni_sub$delivery_ind])$auc
    pr_rf_oos_tempted[ss,jj] <- pr.curve(predprob_rf[metauni_sub$delivery_ind], 
                                      predprob_rf[!metauni_sub$delivery_ind])$auc.integral
  }
  write.csv(pr_glm_oos_tempted, 
            file=paste0("result/realsim_ecam_pr_oos_glm_", method, ".csv"))
  write.csv(roc_glm_oos_tempted, 
            file=paste0("result/realsim_ecam_roc_oos_glm_", method, ".csv"))
  write.csv(pr_rf_oos_tempted, 
            file=paste0("result/realsim_ecam_pr_oos_rf_", method, ".csv"))
  write.csv(roc_rf_oos_tempted, 
            file=paste0("result/realsim_ecam_roc_oos_rf_", method, ".csv"))
}
```


# FTSVD

## Run FTSVD

Save subject loading and trajectory

```{r run_ftsvd, eval=FALSE}
smooth <- c(1, 0.1, 0.01, 0.005, rep(1e-4, 5))
for (ss in 1:nsim){
  print(ss)
  for (jj in 1:length(nkeep)){
    meta_sub <- read.csv(paste0("simdata/realsim", ss, "_ntime", nkeep[jj], ".csv"), header=T, row.names=1)
    count_sub <- count_tab[rownames(meta_sub),]
    meta_sub$studyid <- as.character(meta_sub$studyid)
    meta_sub$delivery_ind <- meta_sub$delivery=="Vaginal"
    metauni <- unique(meta_sub[,c("studyid", "delivery_ind")])
    subdata <- format_tempted(count_sub, meta_sub$visit, meta_sub$studyid, 
                            threshold=0.95, pseudo=0.5,
                            transform="clr")
    # run ftsvd with all subjects and run permanova test
    svd_sub <- svd_centralize(subdata)
    res_ftsvd <- tempted(svd_sub$datlist, r = npc, 
                           resolution = (nkeep[jj]+16)*2, smooth=1e-4)
    agg_feat <- aggregate_feature(res_ftsvd, NULL, subdata)
    aggfeat_mat <- reshape(agg_feat$metafeature_aggregate[,1:4],
                           idvar=c("subID","timepoint") ,
                           v.names=c("value"), timevar="PC",
                           direction="wide")
    colnames(aggfeat_mat)[1] <- "studyid"
    aggfeat_mat <- merge(aggfeat_mat, metauni, by="studyid")
    aggfeat_mat[,2+1:npc] <- apply(aggfeat_mat[,2+1:npc], 2, scale)
    ftsvd_sub <- as.data.frame(res_ftsvd$A_hat)
    ftsvd_sub$`intercept` <- svd_sub$A_tilde
    ftsvd_sub$studyid <- rownames(ftsvd_sub)
    ftsvd_sub <- merge(metauni, ftsvd_sub)
    write.csv(aggfeat_mat, 
              file=paste0("simresult_ftsvd/ftsvd_traj_sim",ss,"_ntime",nkeep[jj], ".csv"))
    write.csv(ftsvd_sub, 
              file=paste0("simresult_ftsvd/ftsvd_subj_sim",ss,"_ntime",nkeep[jj], ".csv"))
  }
}
```

## Sample-level

### PERMANOVA F-value

```{r ftsvd_sample_Fvalue, eval=FALSE}
method <- "ftsvd"
Fmodel_ftsvd <- matrix(NA, nsim, length(nkeep))
colnames(Fmodel_ftsvd) <- paste0("nsample", nkeep)
for (ss in 1:nsim){
  print(ss)
  for (jj in 1:length(nkeep)){
    fname <- paste0(method,"_traj_sim",ss,"_ntime",nkeep[jj], ".csv")
    aggfeat_mat <- read.csv(file.path(paste0("simresult_", method), fname), row.names=1)
    # calculate PERMANOVA F
    dist_aggft <- vegdist(aggfeat_mat[,2+1:npc], method="euclidean")
    res_perm <- adonis2(dist_aggft ~ aggfeat_mat$delivery)
    Fmodel_ftsvd[ss,jj] <- res_perm$F[1]
  }
}

write.csv(Fmodel_ftsvd, 
          file=paste0("result/realsim_ecam_Fvalue_", method, ".csv"))
```

## Subject-level

### In-sample ROC & PR

-   logistic regression
-   random forest

```{r ftsvd_subj_insample_auc, eval=FALSE}
method <- "ftsvd"
roc_sub_glm_ftsvd <- matrix(NA, nsim, length(nkeep))
colnames(roc_sub_glm_ftsvd) <- paste0("nsample", nkeep)
pr_sub_rf_ftsvd <- roc_sub_rf_ftsvd <- pr_sub_glm_ftsvd <- roc_sub_glm_ftsvd
for (ss in 1:nsim){
  print(ss)
  for (jj in 1:length(nkeep)){
    fname <- paste0(method, "_subj_sim",ss,"_ntime",nkeep[jj], ".csv")
    ftsvd_sub <- read.csv(file.path("simresult_ftsvd", fname), row.names=1)
    # glm
    predprob_glm <- rep(NA, nrow(ftsvd_sub))
    for (ii in 1:nrow(ftsvd_sub)){
      glm_fit <- glm(delivery_ind ~ PC1+PC2,
                   data = ftsvd_sub[-ii,], family = "binomial")
      predprob_glm[ii] <- predict(glm_fit, newdata=ftsvd_sub[ii,], type = "response")
    }
    roc_sub_glm_ftsvd[ss,jj] <- roc.curve(predprob_glm[ftsvd_sub$delivery_ind], 
                                        predprob_glm[!ftsvd_sub$delivery_ind])$auc
    pr_sub_glm_ftsvd[ss,jj] <- pr.curve(predprob_glm[ftsvd_sub$delivery_ind], 
                                      predprob_glm[!ftsvd_sub$delivery_ind])$auc.integral
    # random forest
    set.seed(0)
    rf_fit <- randomForest(as.factor(delivery_ind) ~ PC1+PC2,
                   data = ftsvd_sub, strata=as.factor(delivery_ind))
    predprob_rf <- predict(rf_fit, type = "prob")[,"TRUE"]
    roc_sub_rf_ftsvd[ss,jj] <- roc.curve(predprob_rf[ftsvd_sub$delivery_ind], 
                                        predprob_rf[!ftsvd_sub$delivery_ind])$auc
    pr_sub_rf_ftsvd[ss,jj] <- pr.curve(predprob_rf[ftsvd_sub$delivery_ind], 
                              predprob_rf[!ftsvd_sub$delivery_ind])$auc.integral
  }
}

write.csv(roc_sub_glm_ftsvd, 
          file=paste0("result/realsim_ecam_roc_sub_glm_", method, ".csv"))
write.csv(pr_sub_glm_ftsvd, 
          file=paste0("result/realsim_ecam_pr_sub_glm_", method, ".csv"))
write.csv(roc_sub_rf_ftsvd, 
          file=paste0("result/realsim_ecam_roc_sub_rf_", method, ".csv"))
write.csv(pr_sub_rf_ftsvd, 
          file=paste0("result/realsim_ecam_pr_sub_rf_", method, ".csv"))
```

### Out-of-Sample ROC & PR

-   Logistic regression
-   Random forest

```{r run_ftsvd_oos, eval=FALSE}
method <- "ftsvd"
roc_glm_oos_ftsvd <- matrix(NA, nsim, length(nkeep))
colnames(roc_glm_oos_ftsvd) <- paste0("nsample", nkeep)
pr_rf_oos_ftsvd <- roc_rf_oos_ftsvd <- 
  pr_glm_oos_ftsvd <- roc_glm_oos_ftsvd
smooth <- c(1, 0.1, 0.01, 0.005, rep(1e-4, 5))
for (ss in 1:nsim){
  print(ss)
  for (jj in 1:length(nkeep)){
    meta_sub <- read.csv(paste0("simdata/realsim", ss, "_ntime", nkeep[jj], ".csv"), header=T, row.names=1)
    count_sub <- count_tab[rownames(meta_sub),]
    meta_sub$studyid <- as.character(meta_sub$studyid)
    meta_sub$delivery_ind <- meta_sub$delivery=="Vaginal"
    subdata <- format_tempted(count_sub, meta_sub$visit, meta_sub$studyid, 
                            threshold=0.95, pseudo=0.5, transform="clr")
    metauni_sub <- unique(meta_sub[,c("studyid", "delivery_ind")])
    # leave one out prediction
    predprob_glm <- predprob_rf <- rep(NA, length(subdata))
    
    for (ii in 1:length(subdata)){
      print(ii)
      svd_train <- svd_centralize(subdata[-ii])
      res_train <- tempted(svd_train$datlist, r = npc, 
                           resolution = (nkeep[jj]+16)*2, smooth=1e-4)
      A_test <- est_test_subject(subdata[ii], res_train, svd_train)
      dftrain <- data.frame(y=metauni[-ii,"delivery_ind"], x=res_train$A)
      dftest <- data.frame(y=metauni[ii,"delivery_ind"], x=A_test)
      # logistic regression
      glm_fit <- glm(y ~ ., data = dftrain, family = "binomial")
      predprob_glm[ii] <- predict(glm_fit, newdata=dftest, type = "response")
      # random forest
      rf_fit <- randomForest(as.factor(y) ~ ., data = dftrain,
                             strata=as.factor(y))
      predprob_rf[ii] <- predict(rf_fit, newdata=dftest, type = "prob")[,"TRUE"]
    }
    roc_glm_oos_ftsvd[ss,jj] <- roc.curve(predprob_glm[metauni_sub$delivery_ind], 
                                        predprob_glm[!metauni_sub$delivery_ind])$auc
    pr_glm_oos_ftsvd[ss,jj] <- pr.curve(predprob_glm[metauni_sub$delivery_ind], 
                                      predprob_glm[!metauni_sub$delivery_ind])$auc.integral
    roc_rf_oos_ftsvd[ss,jj] <- roc.curve(predprob_rf[metauni_sub$delivery_ind], 
                                        predprob_rf[!metauni_sub$delivery_ind])$auc
    pr_rf_oos_ftsvd[ss,jj] <- pr.curve(predprob_rf[metauni_sub$delivery_ind], 
                                      predprob_rf[!metauni_sub$delivery_ind])$auc.integral
  }
  write.csv(pr_glm_oos_ftsvd, 
            file=paste0("result/realsim_ecam_pr_oos_glm_", method, ".csv"))
  write.csv(roc_glm_oos_ftsvd, 
            file=paste0("result/realsim_ecam_roc_oos_glm_", method, ".csv"))
  write.csv(pr_rf_oos_ftsvd, 
            file=paste0("result/realsim_ecam_pr_oos_rf_", method, ".csv"))
  write.csv(roc_rf_oos_ftsvd, 
            file=paste0("result/realsim_ecam_roc_oos_rf_", method, ".csv"))
}
```

# CTF

## Run CTF

Save subject distance matrix and trajectory

```{r run_ctf, eval=FALSE}
for (ss in 1:nsim){
  print(ss)
  for (jj in 1:length(nkeep)){
    ntime <- nkeep[jj]
    meta_sub <- read.csv(paste0("simdata/realsim", ss, "_ntime", nkeep[jj], ".csv"), header=T, row.names=1)
    meta_sub$studyid <- as.character(meta_sub$studyid)
    meta_sub$delivery_ind <- meta_sub$delivery=="Vaginal"
    meta_sub <- meta_sub[,c("visit", "studyid")]
    count_sub <- count_tab[rownames(meta_sub),]
    count_sub <- count_sub[,colMeans(count_sub==0)<=0.95]
    py_run_file(file="run_ecam_ctf.py",convert=F)
  }
}
```

## Sample-level

### PERMANOVA F-value

```{r ctf_sample_Fvalue, eval=FALSE}
Fmodel_ctf <- matrix(NA, nsim, length(nkeep))
colnames(Fmodel_ctf) <- paste0("nsample", nkeep)
for (ss in 1:nsim){
  print(ss)
  for (jj in 1:length(nkeep)){
    ntime <- nkeep[jj]
    meta_sub <- read.csv(paste0("simdata/realsim", ss, "_ntime", nkeep[jj], ".csv"), header=T, row.names=1)
    count_sub <- count_tab[rownames(meta_sub),]
    meta_sub$studyid <- as.character(meta_sub$studyid)
    meta_sub$delivery_ind <- meta_sub$delivery=="Vaginal"
    metauni <- unique(meta_sub[,c("studyid", "delivery_ind")])
    # calculate PERMANOVA F
    fname <- paste0("distance-matrix_sim", ss, "_ntime", ntime, ".qza")
    ctf_dist <- as.matrix(read_qza(file.path("simresult_ctf", fname))$data)
    deliv <- meta_sub[rownames(ctf_dist),]$delivery
    res_perm <- adonis2(ctf_dist ~ deliv)
    Fmodel_ctf[ss,jj] <- res_perm$F[1]
  }
}

write.csv(Fmodel_ctf, 
          file="result/realsim_ecam_Fvalue_ctf.csv")
```

## Subject-level

### In-sample ROC & PR

-   logistic regression
-   random forest

```{r ctf_subj_insample_glm_auc, eval=FALSE}
roc_sub_glm_ctf <- matrix(NA, nsim, length(nkeep))
colnames(roc_sub_glm_ctf) <- paste0("nsample", nkeep)
pr_sub_rf_ctf <- roc_sub_rf_ctf <- pr_sub_glm_ctf <- roc_sub_glm_ctf
for (jj in 1:length(nkeep)){
  print(jj)
  ntime <- nkeep[jj]
  for (ss in 1:nsim){
    meta_sub <- read.csv(paste0("simdata/realsim", ss, "_ntime", nkeep[jj], ".csv"), header=T, row.names=1)
    count_sub <- count_tab[rownames(meta_sub),]
    meta_sub$studyid <- as.character(meta_sub$studyid)
    meta_sub$delivery_ind <- meta_sub$delivery=="Vaginal"
    metauni_sub <- unique(meta_sub[,c("studyid", "delivery_ind")])
    rownames(metauni_sub) <- metauni_sub$studyid
    fname <- paste0("subject-biplot_sim", ss, "_ntime", ntime, ".qza")
    ctf_sub <- read_qza(file.path("simresult_ctf", fname))$data$Vectors
    colnames(ctf_sub)[1] <- "studyid"
    ctf_sub <- merge(ctf_sub, metauni_sub)
    # logistic regression
    predprob_glm <- rep(NA, nrow(ctf_sub))
    for(ii in 1:nrow(ctf_sub)){
      glm_fit <- glm(delivery_ind~PC1+PC2, data=ctf_sub[-ii,])
      predprob_glm[ii] <- predict(glm_fit, newdata=ctf_sub[ii,], type = "response")
    }
    roc_sub_glm_ctf[ss,jj] <- roc.curve(predprob_glm[ctf_sub$delivery_ind], 
                                    predprob_glm[!ctf_sub$delivery_ind])$auc
    pr_sub_glm_ctf[ss,jj] <- pr.curve(predprob_glm[ctf_sub$delivery_ind], 
                                  predprob_glm[!ctf_sub$delivery_ind])$auc.integral
    # random forest
    set.seed(0)
    rf_fit <- randomForest(as.factor(delivery_ind)~PC1+PC2, data=ctf_sub, strata=as.factor(delivery_ind))
    predprob_rf <- predict(rf_fit, type = "prob")[,"TRUE"]
    roc_sub_rf_ctf[ss,jj] <- roc.curve(predprob_rf[ctf_sub$delivery_ind], 
                                    predprob_rf[!ctf_sub$delivery_ind])$auc
    pr_sub_rf_ctf[ss,jj] <- pr.curve(predprob_rf[ctf_sub$delivery_ind], 
                                  predprob_rf[!ctf_sub$delivery_ind])$auc.integral
  }
}
write.csv(roc_sub_glm_ctf, 
          file="result/realsim_ecam_roc_sub_glm_ctf.csv")
write.csv(pr_sub_glm_ctf, 
          file="result/realsim_ecam_pr_sub_glm_ctf.csv")
write.csv(roc_sub_rf_ctf, 
          file="result/realsim_ecam_roc_sub_rf_ctf.csv")
write.csv(pr_sub_rf_ctf, 
          file="result/realsim_ecam_pr_sub_rf_ctf.csv")
```


# microTensor

## Run microTensor

Save subject loading and trajectory

```{r run_microtensor, eval=FALSE}
set.seed(0)
for (ss in 1:nsim){
  print(ss)
  for (jj in 1:length(nkeep)){
    ntime <- nkeep[jj]
    meta_sub <- read.csv(paste0("simdata/realsim", ss, "_ntime", nkeep[jj], ".csv"), header=T, row.names=1)
    meta_sub$studyid <- as.character(meta_sub$studyid)
    meta_sub$delivery_ind <- meta_sub$delivery=="Vaginal"
    meta_sub$sampleid <- rownames(meta_sub)
    metauni <- unique(meta_sub[,c("studyid", "delivery_ind")])
    count_sub <- count_tab[meta_sub$sampleid,]
    count_sub <- count_sub[,colMeans(count_sub==0)<=0.95]

    X_array <- array(NA, dim = c(ncol(count_sub),
                             length(unique(meta_sub$studyid)),
                             ntime))
    dimnames(X_array) <- list(colnames(count_sub),
                              unique(meta_sub$studyid),
                              1:ntime)
    for(k in seq(1, dim(X_array)[3])) {
      k_df_samples <- meta_sub %>% 
        dplyr::filter(visit == k)
      k_df_samples <- k_df_samples[!duplicated(k_df_samples$studyid),]
      X_array[, k_df_samples$studyid, k] <- 
        t(count_sub[rownames(k_df_samples), ])
    }
    set.seed(0)
    fit_microTensor <- 
      microTensor::microTensor(X = X_array, R = npc, 
                               nn_t = TRUE, ortho_m = TRUE,
                               weighted = TRUE)
    micro_sub <- as.data.frame(fit_microTensor$s)
    colnames(micro_sub) <- paste0("PC",1:npc)
    micro_sub$studyid <- dimnames(X_array)[[2]]
    micro_sub <- merge(metauni, micro_sub)
    write.csv(micro_sub, 
              file=paste0("simresult_micro/micro_subj_sim", ss, "_ntime", ntime, ".csv"))
    micro_traj <- microTensor::create_loading(fit_decomp = fit_microTensor,
                                              feature_names = dimnames(X_array)[[1]],
                                              subject_names = dimnames(X_array)[[2]],
                                              time_names = 1:ntime,
                                              class = "sample")
    colnames(micro_traj) <- c("studyid", "visit", paste0("PC",1:npc))
    micro_traj <- merge(meta_sub[,c("sampleid","day_of_life","studyid", "visit", "delivery_ind")], micro_traj)
    write.csv(micro_traj, 
              file=paste0("simresult_micro/micro_traj_sim", ss, "_ntime", ntime, ".csv"))
  }
}
```


## Sample-level

### PERMANOVA F-value

```{r micro_sample_Fvalue, eval=FALSE}
Fmodel_micro <- matrix(NA, nsim, length(nkeep))
colnames(Fmodel_micro) <- paste0("nsample", nkeep)
for (ss in 1:nsim){
  print(ss)
  for (jj in 1:length(nkeep)){
    fname <- paste0("micro_traj_sim",ss,"_ntime",nkeep[jj], ".csv")
    aggfeat_mat <- read.csv(file.path("simresult_micro", fname), row.names=1)
    # calculate PERMANOVA F
    dist_aggft <- vegdist(aggfeat_mat[,paste0("PC",1:npc)], method="euclidean")
    res_perm <- adonis2(dist_aggft ~ aggfeat_mat$delivery_ind)
    Fmodel_micro[ss,jj] <- res_perm$F[1]
  }
}

write.csv(Fmodel_micro, 
          file=paste0("result/realsim_ecam_Fvalue_micro.csv"))
```

## Subject-level

### In-sample ROC & PR

-   logistic regression
-   random forest

```{r micro_subj_insample_auc, eval=FALSE}
method <- "micro"
roc_sub_glm_micro <- matrix(NA, nsim, length(nkeep))
colnames(roc_sub_glm_micro) <- paste0("nsample", nkeep)
pr_sub_rf_micro <- roc_sub_rf_micro <- pr_sub_glm_micro <- roc_sub_glm_micro
for (ss in 1:nsim){
  print(ss)
  for (jj in 1:length(nkeep)){
    fname <- paste0("micro_subj_sim",ss,"_ntime",nkeep[jj], ".csv")
    micro_sub <- read.csv(file.path(paste0("simresult_", method), fname), row.names=1)
    # glm
    predprob_glm <- rep(NA, nrow(micro_sub))
    for (ii in 1:nrow(micro_sub)){
      glm_fit <- glm(delivery_ind ~ PC1+PC2,
                   data = micro_sub[-ii,], family = "binomial")
      predprob_glm[ii] <- predict(glm_fit, newdata=micro_sub[ii,], type = "response")
    }
    roc_sub_glm_micro[ss,jj] <- roc.curve(predprob_glm[micro_sub$delivery_ind], 
                                        predprob_glm[!micro_sub$delivery_ind])$auc
    pr_sub_glm_micro[ss,jj] <- pr.curve(predprob_glm[micro_sub$delivery_ind], 
                                      predprob_glm[!micro_sub$delivery_ind])$auc.integral
    # random forest
    set.seed(0)
    rf_fit <- randomForest(as.factor(delivery_ind) ~ PC1+PC2,
                           data = micro_sub, strata=as.factor(delivery_ind))
    predprob_rf <- predict(rf_fit, type = "prob")[,"TRUE"]
    roc_sub_rf_micro[ss,jj] <- roc.curve(predprob_rf[micro_sub$delivery_ind], 
                                        predprob_rf[!micro_sub$delivery_ind])$auc
    pr_sub_rf_micro[ss,jj] <- pr.curve(predprob_rf[micro_sub$delivery_ind], 
                              predprob_rf[!micro_sub$delivery_ind])$auc.integral
  }
}

write.csv(roc_sub_glm_micro, 
          file=paste0("result/realsim_ecam_roc_sub_glm_", method, ".csv"))
write.csv(pr_sub_glm_micro, 
          file=paste0("result/realsim_ecam_pr_sub_glm_", method, ".csv"))
write.csv(roc_sub_rf_micro, 
          file=paste0("result/realsim_ecam_roc_sub_rf_", method, ".csv"))
write.csv(pr_sub_rf_micro, 
          file=paste0("result/realsim_ecam_pr_sub_rf_", method, ".csv"))
```




# TCAM

## Run TCAM

Save subject loading

```{r run_tcam, eval=FALSE}
for (ss in 1:nsim){
  print(ss)
  for (jj in 1:length(nkeep)){
    meta_sub <- read.csv(paste0("simdata/realsim", ss, "_ntime", nkeep[jj], ".csv"), header=T, row.names=1)
    count_sub <- count_tab[rownames(meta_sub),]
    meta_sub$studyid <- as.character(meta_sub$studyid)
    meta_sub$delivery_ind <- meta_sub$delivery=="Vaginal"
    subdata <- format_tempted(count_sub, meta_sub$visit, meta_sub$studyid, 
                            threshold=0.95, pseudo=0.5,
                            transform="clr")
    tmp <- simplify2array(subdata)
    dat_sub <- as.data.frame(apply(tmp,1,c))
    studyid <- rep(dimnames(tmp)[[3]], each=dim(tmp)[2])
    dat_sub <- cbind(studyid, dat_sub)
    fnameout <- paste0("simresult_tcam/tcam_score_sim", ss, "_ntime", nkeep[jj], ".csv")
    set.seed(0)
    py_run_file(file="run_ecam_tcam.py",convert=F)
  }
}
```

## Subject-level

### In-sample ROC & PR

-   logistic regression
-   random forest

```{r tcam_subj_insample_auc, eval=FALSE}
method <- "tcam"
roc_sub_glm_tcam <- matrix(NA, nsim, length(nkeep))
colnames(roc_sub_glm_tcam) <- paste0("nsample", nkeep)
pr_sub_rf_tcam <- roc_sub_rf_tcam <- pr_sub_glm_tcam <- roc_sub_glm_tcam
for (ss in 1:nsim){
  print(ss)
  for (jj in 1:length(nkeep)){
    fname <- paste0(method, "_score_sim",ss,"_ntime",nkeep[jj], ".csv")
    tcam_sub <- read.csv(file.path("simresult_tcam", fname), sep="\t",
                         row.names=1, header=T)
    colnames(tcam_sub) <- paste0("PC", 1:npc)
    tcam_sub$studyid <- rownames(tcam_sub)
    tcam_sub <- merge(tcam_sub, metauni)
    # glm
    predprob_glm <- rep(NA, nrow(tcam_sub))
    for (ii in 1:nrow(tcam_sub)){
      glm_fit <- glm(delivery_ind ~ PC1+PC2,
                   data = tcam_sub[-ii,], family = "binomial")
      predprob_glm[ii] <- predict(glm_fit, newdata=tcam_sub[ii,], type = "response")
    }
    roc_sub_glm_tcam[ss,jj] <- roc.curve(predprob_glm[tcam_sub$delivery_ind], 
                                        predprob_glm[!tcam_sub$delivery_ind])$auc
    pr_sub_glm_tcam[ss,jj] <- pr.curve(predprob_glm[tcam_sub$delivery_ind], 
                                      predprob_glm[!tcam_sub$delivery_ind])$auc.integral
    # random forest
    rf_fit <- randomForest(as.factor(delivery_ind) ~ PC1+PC2,
                   data = tcam_sub, strata=as.factor(delivery_ind))
    predprob_rf <- predict(rf_fit, type = "prob")[,"TRUE"]
    roc_sub_rf_tcam[ss,jj] <- roc.curve(predprob_rf[tcam_sub$delivery_ind], 
                                        predprob_rf[!tcam_sub$delivery_ind])$auc
    pr_sub_rf_tcam[ss,jj] <- pr.curve(predprob_rf[tcam_sub$delivery_ind], 
                              predprob_rf[!tcam_sub$delivery_ind])$auc.integral
  }
}

write.csv(roc_sub_glm_tcam, 
          file=paste0("result/realsim_ecam_roc_sub_glm_", method, ".csv"))
write.csv(pr_sub_glm_tcam, 
          file=paste0("result/realsim_ecam_pr_sub_glm_", method, ".csv"))
write.csv(roc_sub_rf_tcam, 
          file=paste0("result/realsim_ecam_roc_sub_rf_", method, ".csv"))
write.csv(pr_sub_rf_tcam, 
          file=paste0("result/realsim_ecam_pr_sub_rf_", method, ".csv"))
```

### Out-of-Sample ROC & PR

-   Logistic regression
-   Random forest

```{r run_tcam_oos, eval=FALSE}
method <- "tcam"
roc_glm_oos_tcam <- matrix(NA, nsim, length(nkeep))
colnames(roc_glm_oos_tcam) <- paste0("nsample", nkeep)
pr_rf_oos_tcam <- roc_rf_oos_tcam <- 
  pr_glm_oos_tcam <- roc_glm_oos_tcam
for (ss in 1:nsim){
  print(ss)
  for (jj in 1:length(nkeep)){
    meta_sub <- read.csv(paste0("simdata/realsim", ss, "_ntime", nkeep[jj], ".csv"), header=T, row.names=1)
    count_sub <- count_tab[rownames(meta_sub),]
    meta_sub$studyid <- as.character(meta_sub$studyid)
    meta_sub$delivery_ind <- meta_sub$delivery=="Vaginal"
    subdata <- format_tempted(count_sub, meta_sub$visit, meta_sub$studyid, 
                            threshold=0.95, pseudo=0.5, transform="clr")
    metauni_sub <- unique(meta_sub[,c("studyid", "delivery_ind")])
    # leave one out prediction
    predprob_glm <- predprob_rf <- rep(NA, length(subdata))
    
    for (ii in 1:length(subdata)){
      print(ii)
      subdata_train <- subdata[-ii]
      tmp <- simplify2array(subdata_train)
      dat_train <- as.data.frame(apply(tmp,1,c))
      studyid <- rep(dimnames(tmp)[[3]], each=dim(tmp)[2])
      dat_train <- cbind(studyid, dat_train)
      
      subdata_test <- subdata[ii]
      tmp <- simplify2array(subdata_test)
      dat_test <- as.data.frame(apply(tmp,1,c))
      studyid <- rep(dimnames(tmp)[[3]], each=dim(tmp)[2])
      dat_test <- cbind(studyid, dat_test)
      fnameout <- "simresult_tcam/tcam_test_score"
      
      set.seed(0)
      py_run_file(file="run_ecam_tcam_oos.py",convert=F)
      
      dftrain <- read.csv(paste0(fnameout, "_train.csv"), sep="\t", row.names=1)
      dftest <- read.csv(paste0(fnameout, "_test.csv"), sep="\t", row.names=1)
      colnames(dftrain) <- colnames(dftest) <- paste0("PC", 1:npc)
      dftrain$studyid <- rownames(dftrain)
      dftrain <- merge(dftrain, metauni_sub)
      dftest$studyid <- rownames(dftest)
      dftest <- merge(dftest, metauni_sub)
      
      # logistic regression
      glm_fit <- glm(delivery_ind ~ PC1+PC2, data = dftrain, family = "binomial")
      predprob_glm[ii] <- predict(glm_fit, newdata=dftest, type = "response")
      # random forest
      rf_fit <- randomForest(as.factor(delivery_ind) ~ PC1+PC2, 
                             data = dftrain, strata=as.factor(delivery_ind))
      predprob_rf[ii] <- predict(rf_fit, newdata=dftest, type = "prob")[,"TRUE"]
    }
    roc_glm_oos_tcam[ss,jj] <- roc.curve(predprob_glm[metauni_sub$delivery_ind],
                                      predprob_glm[!metauni_sub$delivery_ind])$auc
    pr_glm_oos_tcam[ss,jj] <- pr.curve(predprob_glm[metauni_sub$delivery_ind], 
                                      predprob_glm[!metauni_sub$delivery_ind])$auc.integral
    roc_rf_oos_tcam[ss,jj] <- roc.curve(predprob_rf[metauni_sub$delivery_ind], 
                                      predprob_rf[!metauni_sub$delivery_ind])$auc
    pr_rf_oos_tcam[ss,jj] <- pr.curve(predprob_rf[metauni_sub$delivery_ind], 
                                      predprob_rf[!metauni_sub$delivery_ind])$auc.integral
  }
  write.csv(pr_glm_oos_tcam, 
            file=paste0("result/realsim_ecam_pr_oos_glm_", method, ".csv"))
  write.csv(roc_glm_oos_tcam, 
            file=paste0("result/realsim_ecam_roc_oos_glm_", method, ".csv"))
  write.csv(pr_rf_oos_tcam, 
            file=paste0("result/realsim_ecam_pr_oos_rf_", method, ".csv"))
  write.csv(roc_rf_oos_tcam, 
            file=paste0("result/realsim_ecam_roc_oos_rf_", method, ".csv"))
}
```



# PCoA

## Sample-level

### PERMANOVA F-value

```{r pcoa_sample_Fvalue, eval=FALSE}
metric_file <- list.files("data", pattern = "distance_matrix")
metric_file
metric_name <- c("Aitchison", "Bray-Curtis", 
                 "gUniFrac-alpha0", "gUniFrac-alpha1", "gUniFrac-alpha5", 
                 "Jaccard", "unweighted_UniFrac", "Weighted_UniFrac")
distmat_all <- vector(mode="list", length(metric_name))
names(distmat_all) <- metric_name
for (ii in 1:length(metric_name)){
  tmp <- read_qza(paste0("data/", metric_file[ii]))$data
  distmat_all[[ii]] <- as.matrix(tmp)
}

Fmodel <- array(0, dim=c(length(metric_name), nsim, length(nkeep)),
                     dimnames=list(metric_name, paste0("sim",1:nsim), paste0("nsample=",nkeep)))
for (ss in 1:nsim){
  print(ss)
  for (jj in 1:length(nkeep)){
    meta_sub <- read.csv(paste0("simdata/realsim", ss, "_ntime", nkeep[jj], ".csv"), header=T, row.names=1)
    count_sub <- count_tab[rownames(meta_sub),]
    meta_sub$studyid <- as.character(meta_sub$studyid)
    meta_sub$delivery_ind <- meta_sub$delivery=="Vaginal"
    ind_sub <- rownames(meta_sub)
    for (ii in 1:length(metric_name)){
      dist_sub <- distmat_all[[ii]][ind_sub,ind_sub]
      res_perm <- adonis2(dist_sub ~ meta_sub$delivery)
      Fmodel[ii,ss,jj] <- res_perm$F[1]
    }
  }
  for (ii in 1:length(metric_name)){
    write.csv(Fmodel[ii,,], 
       file=paste0("result/realsim_ecam_Fvalue_PCoA_", metric_name[ii], ".csv"))
  }
}
```


# Summarize results

## Summarize subject-level ROC and PR

```{r summary_sub_classification}
method  <- c("tempted", "ctf", "micro", "tcam", "ftsvd")
measure <- c("roc", "pr")
classify <- c("glm", "rf")

tab_auc_subj <- NULL

for (ms in measure){
  for (cls in classify){ 
    # in sample
    for (mthd in method){
      fname <- paste0("result/realsim_ecam_", ms, "_sub_", cls, "_", mthd, ".csv")
      tab0 <- read.csv(fname, row.names=1)
      tab0 <- 1-as.matrix(tab0)
      tab_tmp <- data.frame(auc=as.vector(tab0),
                      nsample=as.vector(t(matrix(rep(nkeep,nsim),length(nkeep),nsim))),
                      method=toupper(mthd), measure=toupper(ms), classify=cls,
                      type="In-Sample")
      tab_auc_subj <- rbind(tab_auc_subj, tab_tmp)
    }
    # out of sample
    for (mthd in c("tempted", "tcam", "ftsvd")){
      fname <- paste0("result/realsim_ecam_", ms, "_oos_", cls, "_", mthd, ".csv")
      tab0 <- read.csv(fname, row.names=1)
      tab0 <- 1-as.matrix(tab0)
      tab_tmp <- data.frame(auc=as.vector(tab0),
                      nsample=as.vector(t(matrix(rep(nkeep,nsim),length(nkeep),nsim))),
                      method=toupper(mthd), measure=toupper(ms), classify=cls,
                      type="Out-of-Sample")
      tab_auc_subj <- rbind(tab_auc_subj, tab_tmp)
    }
  }  
}
tab_auc_subj$classify <- gsub("glm", "Logistic Regression", tab_auc_subj$classify)
tab_auc_subj$classify <- gsub("rf", "Random Forest", tab_auc_subj$classify)
tab_auc_subj$method <- gsub("MICROTENSOR", "microTensor", tab_auc_subj$method)
```

### Plot all subject-level PR

```{r plot_sub_auc}
tab_auc_subj$nsample <- as.factor(tab_auc_subj$nsample)
tab1 <- aggregate(auc~nsample+measure+method+classify+type, data=tab_auc_subj, 
                  FUN=mean)
names(tab1)[6] <- "mean"
tab2 <- aggregate(auc~nsample+measure+method+classify+type, data=tab_auc_subj, 
                  FUN=function(x){sd(x)/sqrt(length(x))})
names(tab2)[6] <- "se"
rownames(tab1) <-rownames(tab2) <- NULL
tab_auc_subj_summary <- merge(tab1, tab2)
tab_auc_subj_summary$nsample <- factor(tab_auc_subj_summary$nsample, 
                                     level=as.character(nkeep))

color_method <- c("#33a02c", #green for CTF
                  "#bdbdbd", #violet for FTSVD
                  "#0096FF", #blue for microTensor
                  "#f525dd", #pink for TCAM
                  "#5A5A5A") #gray for TEMPTED

ann_text <- tab_auc_subj_summary[2,]
ann_text$mean <- 0.5-0.015
ann_text$nsample <- 6.5
p_subj_roc <- ggplot(data=dplyr::filter(tab_auc_subj_summary, measure=="ROC"), 
                           aes(x=nsample, y=mean, group=paste0(type,method), color=method)) + 
  geom_line(size=1, position=position_dodge(0.5), aes(linetype=type)) + 
  geom_point(size=2, position=position_dodge(0.5)) +
  geom_errorbar(aes(ymin=mean-2*se, ymax=mean+2*se, width=0.5), 
                position=position_dodge(0.5), size=1) + 
  geom_hline(yintercept=0.5, linetype="dotted", color = "black", size=1) +
  geom_text(data = ann_text,label = "At random", color="black") +
  facet_wrap(.~classify) +
  labs(y="AUC-ROC error", x="Number of time points") + 
  ggtitle("Subject Level") +
  scale_color_manual(values=color_method) +
  theme_bw()

ann_text <- tab_auc_subj_summary[2,]
ann_text$mean <- mean(!metauni$delivery_ind)-0.015
ann_text$nsample <- 6.5
p_subj_pr <- ggplot(data=dplyr::filter(tab_auc_subj_summary, measure=="PR"), 
                            aes(x=nsample, y=mean, group=paste0(type,method), color=method)) + 
  geom_line(size=1, position=position_dodge(0.5), aes(linetype=type)) + 
  geom_point(size=2, position=position_dodge(0.5)) +
  geom_errorbar(aes(ymin=mean-2*se, ymax=mean+2*se, width=0.5), 
                position=position_dodge(0.5), size=1) + 
  geom_hline(yintercept=mean(!metauni$delivery_ind), linetype="dotted", color = "black", size=1) +
  geom_text(data = ann_text,label = "At random", color="black") +
  facet_wrap(.~classify) +
  labs(y="AUC-PR Error", x="Number of Time Points")+
  ggtitle("Subject Level") +
  scale_color_manual(values=color_method) +
  theme_bw()
p_subj_pr
```

## Summarize sample-level PERMANOVA F-values

```{r summary_Fvalue}
metric_name <- c("Bray-Curtis",
                 "unweighted_UniFrac", "Weighted_UniFrac")
# read in results
Fmodel <- array(0, dim=c(length(metric_name)+4, nsim, length(nkeep)),
                dimnames=list(c(metric_name, "TEMPTED", "CTF", "microTensor", "FTSVD"), paste0("sim",1:nsim), paste0("nsample=",nkeep)))
for (ii in 1:length(metric_name)){
  Fmodel[ii,,] <- as.matrix(read.csv(paste0("result/realsim_ecam_Fvalue_PCoA_", metric_name[ii], ".csv"),
                           row.names=1))
}
Fmodel_tempted <- read.csv("result/realsim_ecam_Fvalue_tempted.csv", row.names=1)
Fmodel[ii+1,,] <- as.matrix(Fmodel_tempted)
Fmodel_ctf <- read.csv("result/realsim_ecam_Fvalue_ctf.csv", row.names=1)
Fmodel[ii+2,,] <- as.matrix(Fmodel_ctf)
Fmodel_micro <- read.csv("result/realsim_ecam_Fvalue_micro.csv", row.names=1)
Fmodel[ii+3,,] <- as.matrix(Fmodel_micro)
Fmodel_ftsvd <- read.csv("result/realsim_ecam_Fvalue_ftsvd.csv", row.names=1)
Fmodel[ii+4,,] <- as.matrix(Fmodel_ftsvd)
# make table
tab_Fmodel <- NULL
for (ii in 1:dim(Fmodel)[1]){
  tab_tmp <- data.frame(Fvalue=as.vector(Fmodel[ii,,]),
                        nsample=as.vector(t(matrix(rep(nkeep,nsim),length(nkeep),nsim))),
                        method=dimnames(Fmodel)[[1]][ii])
  tab_Fmodel <- rbind(tab_Fmodel, tab_tmp)
}
tab_Fmodel$method[tab_Fmodel$method=="unweighted_UniFrac"] <- "UniFrac"
tab_Fmodel$method[tab_Fmodel$method=="Weighted_UniFrac"] <- "Weighted UniFrac"
```

### Plot all PERMANOVA F-values

```{r plot_Fvalue}
tab1 <- aggregate(Fvalue~nsample+method, data=tab_Fmodel, 
                  FUN=mean)
names(tab1)[3] <- "mean"
tab2 <- aggregate(Fvalue~nsample+method, data=tab_Fmodel, 
                  FUN=function(x){sd(x)/sqrt(length(x))})
names(tab2)[3] <- "se"
rownames(tab1) <-rownames(tab2) <- NULL
tab_Fmodel_summary <- merge(tab1, tab2)

color_method <- c("#ff7f00", #orange for Bray
                  "#33a02c", #green for CTF
                  "#bdbdbd", #violet for FTSVD
                  "#0096FF", #blue for microTensor
                  "#5A5A5A", #gray for TEMPTED
                  "#7F00FF", #purple for Unifrac
                  "#e31a1c") #red for unweighted Unifrac
p_Fmodel_summary <- ggplot(data=tab_Fmodel_summary, 
                                aes(x=nsample, y=mean, group=method, color=method)) + 
  geom_line(size=1, position=position_dodge(0.1)) + geom_point(size=2, position=position_dodge(0.1)) +
  geom_errorbar(aes(ymin=mean-2*se, ymax=mean+2*se, width=1), 
                position=position_dodge(0.1), size=1) + 
  scale_x_continuous(breaks=2:10) + 
  scale_color_manual(values=color_method) +
  ylab("PERMANOVA F-value") + xlab("Number of Time Points") + 
  ggtitle("Sample Level") +
  theme_bw() + 
  theme(legend.position = "bottom")
p_Fmodel_summary
```


#### get legend color for all

```{r get_legend}
color_method <- c("#ff7f00", #orange for Bray
                  "#33a02c", #green for CTF
                  "#bdbdbd", #violet for FTSVD
                  "#0096FF", #blue for microTensor
                  "#f525dd", #pink for TCAM
                  "#5A5A5A", #gray for TEMPTED
                  "#7F00FF", #purple for Unifrac
                  "#e31a1c") #red for unweighted Unifrac
mm <- length(color_method)
tab_lgd <- data.frame(Method=rep(c("Bray-Curtis", "CTF", "FTSVD", "microTensor", "TCAM",
                               "TEMPTED", "UniFrac","Weighted UniFrac"),4),
                      value=rnorm(4*mm), time=rep(1:4, each=mm),
                      Type=c(rep("In-Sample",each=2*mm), rep("Out-of-Sample", each=2*mm)))
p_lgd <- ggplot(data=tab_lgd, aes(x=time, y=value, color=Method)) + 
  geom_point(size=2) + geom_line(aes(linetype=Type), size=1) + 
  scale_color_manual(values=color_method) +
  theme_bw() +
  theme(legend.position="bottom") +
  guides(color = guide_legend(nrow = 2), linetype = guide_legend(nrow = 2))
p_lgd
lgd <- get_legend(p_lgd)
```

## Print plot for manuscript

```{r plot_all}
lay <- rbind(c(1,1,2),c(1,1,2),c(1,1,2),c(1,1,2),c(1,1,2),
             c(3,3,3))
p_all <- grid.arrange(p_subj_pr + theme(legend.position="none"),
                      p_Fmodel_summary + theme(legend.position="none"), 
                      lgd,
                      layout_matrix=lay)
ggsave("../figure_table/realsim_ecam.pdf", width=10, height=5, plot=p_all)
```
